### 2017-06-06 inst/paper/figs.Rmd
init commit

---
title: "F1000 paper figures"
author: "Kent Riemondy RBI"
date: "June 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_libs}
library(valr)
library(tidyverse)
library(cowplot)
```
## Figure 1

```{r, fig1}

# theme to make pretty for publication

pub_theme <- theme(axis.text = element_text(size = 14),
                   plot.title = element_text(size = 14),
                   strip.text = element_text(size = 14))

x <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1', 25,     50,
  'chr1', 100,    125
)

y <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1', 30,     75
)

a <- bed_glyph(bed_intersect(x, y)) + pub_theme

x <- tibble::tribble(
  ~chrom, ~start, ~end,
  'chr1',      1,      50,
  'chr1',      10,     75,
  'chr1',      100,    120
)

b <- bed_glyph(bed_merge(x)) + pub_theme
plot_grid(a, b, align = 'h', nrow = 1, labels="AUTO", scale = c(.9, .9))
ggsave("glyphs.pdf", width = 8, height = 4)

```

## Figure 2
```{r fig2}
# `valr_example()` identifies the path of example files
bedfile <- valr_example('genes.hg19.chr22.bed.gz')
genomefile <- valr_example('hg19.chrom.sizes.gz')
bgfile  <- valr_example('hela.h3k4.chip.bg.gz')

genes <- read_bed(bedfile, n_fields = 6)
genome <- read_genome(genomefile)
y <- read_bedgraph(bgfile)

# generate 1 bp TSS intervals, `+` strand only
tss <- genes %>%
  filter(strand == '+') %>%
  mutate(end = start + 1)

# 1000 bp up and downstream
region_size <- 1000
# 50 bp windows
win_size <- 50

# add slop to the TSS, break into windows and add a group
x <- tss %>%
  bed_slop(genome, both = region_size) %>%
  bed_makewindows(genome, win_size)

x

# map signals to TSS regions and calculate summary statistics.
res <- bed_map(x, y, win_sum = sum(value, na.rm = TRUE)) %>%
  group_by(.win_id) %>%
  summarize(win_mean = mean(win_sum, na.rm = TRUE),
            win_sd = sd(win_sum, na.rm = TRUE))

res

x_labels <- seq(-region_size, region_size, by = win_size * 5)
x_breaks <- seq(1, 41, by = 5)

sd_limits <- aes(ymax = win_mean + win_sd, ymin = win_mean - win_sd)

p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
  xlab('Position (bp from TSS)') + ylab('Signal') + 
  pub_theme +
  theme_classic()

plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
ggsave("tss.pdf", width = 3.5, height = 2.25)
```

## Figure 3

```{r, message=FALSE}
# Load required packages
library(valr)
library(tidyverse)
library(scales)
library(microbenchmark)
library(stringr)
library(cowplot)

# Set-up test genome
genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
write_tsv(genome, "genome.txt", col_names=FALSE)

# number of intervals
n <- 1e6
# number of timing reps
nrep <- 10

seed_x <- 1010486
x <- bed_random(genome, n = n, seed = seed_x)
write_tsv(x, "x.bed", col_names=FALSE)
seed_y <- 9283019
y <- bed_random(genome, n = n, seed = seed_y)
write_tsv(y, "y.bed", col_names=FALSE)

res <- microbenchmark(
  # randomizing functions
  bed_random(genome, n = n, seed = seed_x),
  bed_shuffle(x, genome, seed = seed_x),
  # # single tbl functions
  bed_slop(x, genome, both = 1000),
  bed_flank(x, genome, both = 1000),
  bed_merge(x),
  bed_cluster(x),
  bed_complement(x, genome),
  # multi tbl functions
  bed_closest(x, y),
  bed_intersect(x, y),
  bed_map(x, y, .n = n()),
  bed_subtract(x, y),
  # stats
  bed_reldist(x, y),
  bed_jaccard(x, y),
  bed_fisher(x, y, genome),
  #utils 
  bed_makewindows(x, genome, win_size = 100),
  times = nrep,
  unit = 's')

load_file <- function(filename) {
  if(endsWith(filename, ".txt")){
    read_genome(filename)
  } else if (endsWith(filename, ".bed")) {
    read_bed(filename)
  }
}

genome_file <- "genome.txt"
x_bed <- "x.bed"
y_bed  <- "y.bed"

res_hard <- microbenchmark(
  # randomizing functions
  bed_random(load_file(genome_file), n = n, seed = seed_x),
  bed_shuffle(load_file(x_bed), load_file(genome_file), seed = seed_x),
  # # single tbl functions
  bed_slop(load_file(x_bed), load_file(genome_file), both = 1000),
  bed_flank(load_file(x_bed), load_file(genome_file), both = 1000),
  bed_merge(load_file(x_bed)),
  bed_cluster(load_file(x_bed)),
  bed_complement(load_file(x_bed), load_file(genome_file)),
  # multi tbl functions
  bed_closest(load_file(x_bed), load_file(y_bed)),
  bed_intersect(load_file(x_bed), load_file(y_bed)),
  bed_map(load_file(x_bed), load_file(y_bed), .n = n()),
  bed_subtract(load_file(x_bed), load_file(y_bed)),
  # stats
  bed_reldist(load_file(x_bed), load_file(y_bed)),
  bed_jaccard(load_file(x_bed), load_file(y_bed)),
  bed_fisher(load_file(x_bed), load_file(y_bed), load_file(genome_file)),
  #utils 
  bed_makewindows(load_file(x_bed), load_file(genome_file), win_size = 100),
  times = nrep,
  unit = 's')

# covert nanoseconds to seconds
res <- res %>%
  as_tibble() %>%
  mutate(time = time / 1e9)

res_hard <- res_hard %>%
  as_tibble() %>%
  mutate(time = time / 1e9)

# futz with the x-axis
sts <- boxplot.stats(res$time)$stats
# filter out outliers (currently filtering fisher and map)
# res <- filter(res, time <= max(sts) * 1.05)

```

```{bash}
# Function to benchmark any bash command
repeats=10

bench_tests() {
    # --------------------------------------------------------------------------
    # Benchmark loop
    # --------------------------------------------------------------------------
    echo 'Benchmarking ' $1'...';

    # Run the given command [repeats] times
    for (( i = 1; i <= $repeats ; i++ ))
    do
        # Indicate the command we just ran in the csv file
        echo -ne $3"\t" >> $2
        # runs time function for the called script, output in a comma seperated
        # format output file specified with -o command and -a specifies append
        # requires gnu-time not BSD
        gtime -f %e -o ${2} -a ${1} > /dev/null 2>&1
    done;

}

# bedtools time
# set-up files for bedtools function
bedtools sort -i y.bed > y1.bed
bedtools sort -i x.bed > x1.bed
sort -k1,1 genome.txt > genome1.txt

## Run timing test
rm -f time.txt

bench_tests "bedtools random -g genome1.txt -n 1000000 -seed 1010486" "time.txt" "random"
bench_tests "bedtools shuffle -i x.bed -g genome1.txt -seed 1010486" "time.txt" "shuffle"
# single tbl function
bench_tests "bedtools slop -i x1.bed -g genome1.txt -b 1000" "time.txt" "slop"
bench_tests "bedtools flank -i x1.bed -g genome1.txt -b 1000" "time.txt" "flank"
bench_tests "bedtools merge -i x1.bed" "time.txt" "merge"
bench_tests "bedtools cluster -i x1.bed" "time.txt" "cluster"
bench_tests "bedtools complement -i x1.bed -g genome1.txt" "time.txt" "complement"
# multi tbl functions
bench_tests "bedtools closest -a x1.bed -b y1.bed" "time.txt" "closest"
bench_tests "bedtools intersect -a x1.bed -b y1.bed" "time.txt" "intersect"
bench_tests "bedtools map -a x1.bed -b y1.bed -c 1 -o count" "time.txt" "map"
bench_tests "bedtools subtract -a x1.bed -b y1.bed" "time.txt" "subtract"
# stats
bench_tests "bedtools reldist -a x1.bed -b y1.bed" "time.txt" "reldist"
bench_tests "bedtools jaccard -a x1.bed -b y1.bed" "time.txt" "jaccard"
bench_tests "bedtools fisher -a x1.bed -b y1.bed -g genome1.txt" "time.txt" "fisher"
#utils 
bench_tests "bedtools makewindows -b x1.bed -w 100" "time.txt" "makewindows"
```

```{r}
# Code to process data from previous chunks (above)

# This is to import the data from bedtools (bash) and graph it
times <- read_tsv("time.txt", col_names = c("expr", "time"))
dat <- list(times, res_hard)
names(dat) <- c("BEDTools", "valr")
dat <- bind_rows(dat, .id = "package")

mean_dat <- group_by(dat, expr, package) %>% 
             summarize(mean = mean(time), se = sd(time))

mean_dat$func <- ifelse(mean_dat$package == "bedtools", 
                        mean_dat$expr, 
                        str_split(mean_dat$expr, "\\(",  simplify = T)[, 1] %>% 
  str_split(., "_", simplify = T) %>% 
  .[,2])

ggplot(mean_dat, aes(x=reorder(func, mean), y = mean, fill = package)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
                width = .1, position = position_dodge(1)) +
  scale_fill_brewer(palette = "Paired") + 
  coord_flip() +
  theme_bw() +
  labs(
    y='execution time (seconds)',
    x='',
    subtitle = paste(comma(n), "random 1 kilobase intervals,", 
                     comma(nrep), "repetitions")) +
  theme_classic() +
  pub_theme +
  theme(
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14)
  )

```